This code allows for R to connect to a PostgreSQL database. 
There are steps prior to running this code:
1) Connect to the VPN if working remotely
2) Use the Windows search bar to find the "ODBC Data Source Administrator 64-bit" application
3) Under "User DSN" tab, click the button to "Add..." a connection
4) Choose the connection type that your database is
5) Enter the connection parameters, username, and password for databases you will be using. You either already have this for use with DB visualizer/NaviCat/etc or can obtain from the person who manages databases.
    Data Source = Give it a name of your choice
    Database = the name of the database as you currently have it named when used through other connections
    Description = Give it a description of your choice
    Server = find this in the database configurations
    Port = find this in the database configurations
    User Name = the login Name for the database
    Password = the password for the database

This video is helpful, but choose the type you need as the connection type
"https://www.youtube.com/watch?v=PoA23UWvXuc#"

# Load the package to make connections
```{r}
library(RODBC)
```
    
# Connect to the database name that you created in Windows ODBC 64-bit application
```{r}
myConnection <- RODBC::odbcConnect("databaseName", rows_at_time=1)
```

# Save the sql query I will need, as text
## use the format schema.table when selecting!
```{r}
sql <- "
select * from test.test
limit 5
"
```

# Pull in data and make a table!
```{r}
myData <- sqlQuery(myConnection, sql, stringsAsFactors=FALSE)
```

# Make a new view in the database! -CUSTOMIZE THESE PARAMETERS
```{r}
sqlSave(channel, dat, tablename = NULL, append = FALSE,
        rownames = TRUE, colnames = FALSE, verbose = FALSE,
        safer = TRUE, addPK = FALSE, typeInfo, varTypes,
        fast = TRUE, test = FALSE, nastring = NULL)
```

# Update a view in the database!
```{r}
sqlUpdate(channel, dat, tablename = NULL, index = NULL,
          verbose = FALSE, test = FALSE, nastring = NULL,
          fast = TRUE)
```

See more: 
https://rdrr.io/cran/RODBC/man/sqlSave.html
