This code takes in HL7 reports either as a single or multiple, and:
1) scrapes text
2) separates sections of the report
3) summarizes error types and error descriptions
4) cleans up the categories and text descriptions
5) plots the counts of specific errors
6) creates summary tables
7) creates patient line item spreadsheets by error
8) looks for successful transmissions by date and patient

The report dates are captured for each error

```{r}
library(pdftools)
library(svDialogs)
library(stringr)
library(tidyverse)
library(ggplot2)
library(openxlsx)
library(anytime)
library(lubridate)
```

Options for loading entire year or daily HL7 reports
```{r}
loadHow <- dlg_input("What type of PDF do you have? Enter \n1 for a single PDF \n2 for multiple daily PDFs (one date per PDF)")$res

if(loadHow=="1"){# Load the master PDF from the HL7 reports
  ## EACH PAGE IS ONE ELEMENT ONCE LOADED
  dlg_message("Please select your HL7 file", type ="ok")
  as_text <- pdftools::pdf_text(dlg_open()$res) # text only
  
  # See what it looks like
  as_text[1]
  as_text[35]
}
if(loadHow=="2"){
  dlg_message("Please select the folder containing your HL7 files", type ="ok")
  # Where is the main folder located?
  location <- dlgDir()$res
  # Find the file paths that match your pattern
  subfolders <- dlg_message("Would you like to include subfolders?", type="yesno")$res
  if(subfolders=="yes"){subfolders<-TRUE}else(subfolders<-FALSE)
  my_paths <- list.files(location, pattern = "HL7 Report", recursive = TRUE, include.dirs = paste0(subfolders))
  # Load the files
  loaded <-  lapply(paste0(location, "/", my_paths), function(x)pdftools::pdf_text(x))
  # Combine all the data by starting with the first PDF and appending all others
  as_text <- loaded[[1]]
  for(i in loaded[2:length(loaded)]){
    as_text <- append(as_text, i)
  }
}
```

Subset the PDF pages by type
```{r}
# Find each page by the title, use these page title locations to subset the data
Summary <- as_text[grep("HL7 Transfer Report", as_text)]
Errors <- as_text[grep("Errors from Transferred Data", as_text)]
Users <- as_text[grep("User Errors", as_text)]
Vim <- as_text[grep("Transferred Vim Transactions", as_text)]
```

    
    ----WORK WITH THE TRANSFER ERRORS----

Pull the headers with total and type of error
```{r}
# Set up an empty table to capture all error categories
error_counts <- tibble("Error" = "NA")

# Find every type of error in the PDF, along with the total count for that type
for(i in 1:length(Errors)){
  found <- str_extract_all(Errors[i], "Error Count\\\n\\D+\\d+", simplify = TRUE) # extract string starting with Error Count and ending with the count
  for(j in found){ # save each type and count on each page (page may have more than one type of error listed)
    error_counts <- rbind(error_counts, j)
  }
}

# Remove the empty rows (from setup of table)
error_counts <- error_counts[2:dim(error_counts)[1],]

# Remove extra text by writing and applying a function to substitute text
f <- function(x){gsub("Error Count\n", "", x)}
error_counts <- sapply(error_counts[,1], f)

# Split the type and the count
error_summary <- as_tibble(str_split(error_counts,  "\\W{2,}", simplify = TRUE))
colnames(error_summary) <- c("Error_Type", "Total Count")

# Summarize - combining any repeat types from multiple PDFs
error_summary <- error_summary %>% group_by(Error_Type) %>% summarise("Total Count" = sum(as.numeric(`Total Count`)))
```

Start a patient error record table
```{r}
# Get each occurrance of the error category in one column
records <- "NA"
for(i in error_summary$Error_Type){
  repeats <- rep(i, as.numeric(error_summary$`Total Count`[error_summary$Error_Type==i]))
  for(j in repeats){
    records <- rbind(records, j)
  }
}
# clean up the table
records <- tibble("Error_Category" = records)
records <- records[-1,]
# make empty fields ready
records$Record_Date_Start <- ""
records$Record_Date_End <- ""
records$Patient <- ""
records$Error_Desc <- ""
```

Parse the patient info and errors by description
Add to the records table
```{r}
# Find locations of patient name and DOB strings
## The DOB and name occur between newlines, with DOB in middle, but newline character is not reliable on right side
## Name format is not reliable, so use .+ for left side
## Use newline to find beginning, then use DOB format to find end
DOB_where <- str_locate_all(Errors, "\\n.+(DOB:)\\s+\\d+\\/\\d+\\/\\d+")

# Make sure all DOB are found
a <- 0
for(i in DOB_where){a <- a+(dim(i)[1])}
if(a!=sum(as.numeric(error_summary$`Total Count`))){stop("Hmm...you seem to have missed some records. Please check your DOB regex compared to the PDF. \n(More or fewer records found than are in the summary table)")}

# Pull patient info and error description for each error
n <- 1 # line in the final table
for(i in 1:length(Errors)){
  # Grab the date for this report
    # nested pull the right date, remove surrounding text, separate date range
  page_date_start <- gsub("\\s|(.*-)", "", 
                          gsub("(\\n)|(Data submission date range:)|(Interface)", "", str_extract(Errors[i], "\\nData submission date range:.*\\nInterface")))
  page_date_end <- gsub("\\s|(-.*)", "", 
                        gsub("(\\n)|(Data submission date range:)|(Interface)", "", str_extract(Errors[i], "\\nData submission date range:.*\\nInterface")))
  # if the user loaded more than one PDF ("daily" option) check that the date range is one day only
    ## If not, stop the code and ask for review. 
  if(loadHow=="2" & (difftime(anydate(page_date_end), anydate(page_date_end), units = "days")>1)){
  stop(paste("Page", i, "of your Errors data is a PDF page that does not follow the 'daily' format. Make sure each 'daily' PDF represents a 24 hour period - correct before continuing."))}
  if(length(DOB_where[[i]]==0)){} # If there are no errors on this page, skip it
  if(length(DOB_where[[i]]>0)){
    for(j in 1:dim(DOB_where[[i]])[1]){
    # Use the locations found with str_locate above
    start <- DOB_where[[i]][j,1]
    end <- DOB_where[[i]][j,2]
    x <- substr(Errors[i], start, end) # Pull the patient info
    # Find the description of error for this patient (after DOB)
    xx <- substr(Errors[i], end+1, end+100)
    # Add to the table
    records$Patient[n] <- x 
    records$Error_Desc[n] <- xx
    records$Record_Date_Start[n] <- page_date_start
    records$Record_Date_End[n] <- page_date_end
    # Count up for next row in table
    n <- n+1
    }
  }
}

# review if any descriptions need to be added to the regex
t <- table(records$Error_Desc)
view(t)
dlg_message("Please review the error descriptions. If there are any not represented in the regex for the next block mutate step, add before continuing", type ="ok")
```

Clean up the table
```{r}
if(dim(records)[1]==0){stop("You have no Transfer Errors, skipping.")} # Skip if empty

# Separate Name and DOB
records$DOB <- str_extract(records$Patient, "\\d+\\/\\d+\\/\\d+")
records$Patient <- str_extract(records$Patient, "\\w+,\\W+\\w+")

# Clean up certain Error Types that do not have descriptions
records$Error_Desc[grepl("(More Than One)|(Person Is Deceased)", records$Error_Category, ignore.case = TRUE)] <- "NA"

# Standardize the error description
  ## IF absent, will save category
records <- records %>% mutate(
  Error_Desc = case_when(
    grepl("Vaccine reported in inventory for lot number", Error_Desc, ignore.case = TRUE) ~ "Wrong lot number", 
    grepl("No record to delete", Error_Desc, ignore.case = TRUE) ~ "No record to delete", 
    grepl("Unable to delete dose.+More than one", Error_Desc, ignore.case = TRUE) ~ "Can't delete duplicate shot with conflicting lot",
    grepl("waiver record already e", Error_Desc, ignore.case = TRUE) ~ "Waiver record already exists for series",
    grepl("Shot Date cannot be less than birth", Error_Desc, ignore.case = TRUE) ~ "Shot date prior to DOB",
    grepl("Manufacturer \\w{3} not found for vaccine", Error_Desc, ignore.case = TRUE) ~ "Invalid vaccine and manufacturer match",
    grepl("Failed VIM Transactions", Error_Desc, ignore.case = TRUE) ~ "Vaccine inventory not deducted",
    grepl("This site already has p", Error_Desc, ignore.case = TRUE) ~ "Conflicting info in existing record",
    grepl("Child not found", Error_Desc, ignore.case = TRUE) ~ "Cannot delete - not in MCIR",
    grepl("area code", Error_Desc, ignore.case = TRUE) ~ "Incorrect area code",
    grepl("Responsible part \\w+ must be", Error_Desc, ignore.case = TRUE) ~ "Issue with responsible party info",
    grepl("numeric or value error: character", Error_Desc, ignore.case = TRUE) ~ "Incorrect ZIP code format",
    grepl("p_process_patient Error", Error_Desc, ignore.case = TRUE) ~ "Vaccine did not process",
    grepl("Vaccine \\d+ .* is not in tab", Error_Desc, ignore.case = TRUE) ~ "Invalid vaccine code",
    grepl("Lot not found", Error_Desc, ignore.case = TRUE) ~ "Vaccine lot not in inventory",
    grepl("site already has this pat", Error_Desc, ignore.case = TRUE) ~ "Patient id already exists", 
    grepl("exact fetch returns more than requested", Error_Desc, ignore.case = TRUE) ~ "Fetch row error", 
    grepl("error returned from p_post_child_race", Error_Desc, ignore.case = TRUE) ~ "Cannot update race", 
    grepl("Residence state code \\w{2,3} not in table", Error_Desc, ignore.case = TRUE) ~ "State code not found",
    grepl("returned from p_update_resp_party", Error_Desc, ignore.case = TRUE) ~ "Cannot update resp_party",
    grepl("Birth Date Cannot be more than child's shot", Error_Desc, ignore.case = TRUE) ~ "Child DOB after shot date", 
    grepl("UTILITIES.F_CHECK_PRIV:", Error_Desc, ignore.case = TRUE) ~ "Utilities Check", 
    TRUE ~ Error_Category
  )
)
table(records$Error_Desc)
```


Summarize and print the transfer errors
```{r}
if(dim(records)[1]==0){stop("You have no Transfer Errors, skipping.")} # Skip if empty

summary_with_desc <- records %>% group_by(Error_Category, Error_Desc) %>% summarise(n = n())

desc_plot <- ggplot(records) +
  aes(x = Error_Desc, fill = Error_Desc) +
  geom_bar(position = "dodge") +
  labs(
    title = "HL7 Transfer Errors",
    x = "",
    y = "Number of errors",
    fill = "Error description"
  ) +
  theme_bw() + 
  geom_text(data=summary_with_desc,aes(x=Error_Desc,y=n,label=n),vjust=0.5, hjust=-0.25) +
  coord_flip() +
  scale_y_log10() +
  theme(legend.position = "none") +
  expand_limits(y=2500)

desc_plot

dlg_message("Select the folder to save your output")
DIR <- dlg_dir()$res
write.xlsx(summary_with_desc, paste0(DIR, "/Summary of transfer error descriptions.xlsx"), overwrite = TRUE)
write.xlsx(records, paste0(DIR, "/Records with tranfer errors.xlsx"), overwrite = TRUE)

```


    ----WORK WITH THE USER ERRORS----
    
Strip the header
```{r}
# set up the match object using a pattern and the text
pat <- regexpr("\\s+User Errors Report\\n.+\\n.*{5}\\d+\\/\\d+\\/\\d+\\n", Users) 
# Apply the pattern object and find all matches (One for each page)
user_header <- regmatches(Users, pat)

# Strip the headers from the pages, and replace with 2 newline characters to match the formatting at the beginning of all other records in this page
## (every patient record starts with \n\n except the first one, and we stripped one with the header - this will fix that)
Users <- gsub("\\s+User Errors Report\\n.+\\n.*{5}\\d+\\/\\d+\\/\\d+\\n", "\n\n", Users)
```


```{r}
# Find locations of patient name patient id, and DOB strings
## The names occur after two newlines, and are the first item in each record. They end with whitespace and then the title "Original Value"
## Name format is not reliable, so allow for whitespace and other characters, and don't rely on double newline
## DOB occurs after the title "Status" and a newline, then "DOB:" precedes. 
## Patient id follows DOB and always has "PATIENT ID:" preceding
DOB_where <- str_locate_all(Users, "(DOB:)\\s+(\\d+\\/\\d+\\/\\d+)")
Name_where <- str_locate_all(Users, "\\n.+,.+Original")
Id_where <- str_locate_all(Users, "PATIENT ID:.*\\n")

# Make sure all the records are found (each record should have name, DOB, and id)
a <- 0
b <- 0
c <- 0
for(i in DOB_where){a <- a+(dim(i)[1])}
for(i in Name_where){b <- b+(dim(i)[1])}
for(i in Id_where){c <- c+(dim(i)[1])}
# Make sure all three lengths are the same - same number of records found 3 ways
if(!((a==b)&(b==c))){stop("Hmm...you seem to have missed some records. Please check your DOB regex compared to the PDF. \n(More or fewer records found than are in the summary table)")}
```


Parse the patient info and errors by description
```{r}
# Set up a table to hold the errors and records
user_error_records <- tibble(
  "Record_Date_Start" = rep("NA", a), 
  "Record_Date_End" = rep("NA", a),
  "Name" = rep("NA", a),
  "DOB" = rep("NA", a), 
  "Patient Id" = rep("NA", a),
  "Error" = rep("NA", a))

# Pull patient info and error description for each error
n <- 1
nn <- 1
nnn <- 1
for(i in 1:length(Users)){ # for each page of user errors
    page_date_start <- gsub("\\s|(.*-)", "", 
                            gsub("(\\n)|(Data submission date range:)|(Interface)", "", str_extract(Errors[i], "\\nData submission date range:.*\\nInterface")))
    page_date_end <- gsub("\\s|(-.*)", "", 
                          gsub("(\\n)|(Data submission date range:)|(Interface)", "", str_extract(Errors[i], "\\nData submission date range:.*\\nInterface")))  
  if(length(DOB_where[[i]]==0)){} # If there are no errors on this page, skip it
  if(length(DOB_where[[i]]>0)){
    for(j in 1:dim(DOB_where[[i]])[1]){ # for each dob in that page
      # pull start and end index for this item
      DOB_start <- DOB_where[[i]][j,1]
      DOB_end <- DOB_where[[i]][j,2]
      # use the indices to grab the string
      D <- substr(Users[i], DOB_start, DOB_end)
      # apply the string to the records table
      user_error_records$DOB[n] <- D
      n <- n+1 # iterate up for next loop & next row in final table
      } 
    for(k in 1:dim(Name_where[[i]])[1]){ # for each name in that page (same logic as for DOB)
      Name_start <- Name_where[[i]][k,1]
      Name_end <- Name_where[[i]][k,2]
      Nm <- substr(Users[i], Name_start, Name_end)
      user_error_records$Name[nn] <- Nm
      nn <- nn+1 # iterate up for next loop & next row in final table
    }
    for(m in 1:dim(Id_where[[i]])[1]){
      Id_start <- Id_where[[i]][m,1]
      Id_end <- Id_where[[i]][m,2]
      I <- substr(Users[i], Id_start, Id_end)
      user_error_records$`Patient Id`[nnn] <- I
      # Pull the error message from after the patient id
      E <- str_extract(substr(Users[i], Id_end, Id_end+300), "\\n.*\\n")
      user_error_records$Error[nnn] <- E
      # get the dates for this record page
      user_error_records$Record_Date_Start[nnn] <- page_date_start
      user_error_records$Record_Date_End[nnn] <- page_date_end
      nnn <- nnn+1 # iterate up for next loop & next row in final table
    }
  }
}
```
    
Clean up the user error table
```{r}
if(dim(user_error_records)[1]==0){stop("You have no User Errors, skipping.")} # Skip if empty

# Get rid of extra info in the fields
user_error_records$Name <- gsub("(\\n)|(\\s+Original)", "", user_error_records$Name)
user_error_records$DOB <- gsub("\\s|(DOB:)", "", user_error_records$DOB)
user_error_records$`Patient Id` <- gsub("(PATIENT ID:)|\\n", "", user_error_records$`Patient Id`)

# Clean up the error descriptions - make into two fields
split_error <- str_split(user_error_records$Error, "\\s{5,}.*\\s{5,}", simplify = TRUE) # split on the garbage in the middle of the string
colnames(split_error) <- c("Error_Type", "Error_Description")
# Remove newlines
split_error <- gsub("\\n", "", split_error)
# Bind clean fields to final data and remove dirty error field
user_error_records <- cbind(user_error_records, split_error) 
user_error_records <- user_error_records %>% select(-Error)
```

Clean up the error descriptions
```{r}
if(dim(user_error_records)[1]==0){stop("You have no User Errors, skipping.")} # Skip if empty
user_error_records <- user_error_records%>% mutate(
  Error_Description = case_when(
    grepl("phone area code.*must be 3 numeric", Error_Description, ignore.case = TRUE) ~ "Area code not 3 digits", 
    grepl("Gender : code", Error_Description, ignore.case = TRUE) ~ "Bad gender code", 
    grepl("(should be all letters)|(should be valid alphabet)", Error_Description, ignore.case = TRUE) ~ "Should be all letters", 
    grepl("Manufacturer : code", Error_Description, ignore.case = TRUE) ~ "Bad manufacturer code", 
    grepl("The Date cannot be greater than today", Error_Description, ignore.case = TRUE) ~ "Date is the in future", 
    grepl("phone number.*must be 7", Error_Description, ignore.case = TRUE) ~ "Phone number not 7 digits", 
    grepl("State : code", Error_Description, ignore.case = TRUE) ~ "Bad state code", 
    grepl("must be present to add a child", Error_Description, ignore.case = TRUE) ~ "Missing info can't add child", 
    grepl("Shot Date cannot be less than birth", Error_Description, ignore.case = TRUE) ~ "Shot date prior to DOB", 
    grepl("Vaccine_Num : code", Error_Description, ignore.case = TRUE) ~ "Bad Vaccine_Num code", 
    grepl("PRF not found for vaccine code", Error_Description, ignore.case = TRUE) ~ "Bad vaccine code", 
    grepl("Phone Area code can not be", Error_Description, ignore.case = TRUE) ~ "Missing area code", 
    TRUE & !is.na(Error_Description) & Error_Description!="NA" ~ Error_Description,
    TRUE ~ Error_Type
  )
)
t <- table(user_error_records$Error_Description)
view(t)
```

Summarize and print the user errors
```{r}
if(dim(user_error_records)[1]==0){stop("You have no User Errors, skipping.")} # Skip if empty
# Summarize the errors
summary_with_desc2 <- user_error_records %>% group_by(Error_Type, Error_Description) %>% summarise(n = n())

# plot
desc_plot2 <- ggplot(user_error_records) +
  aes(x = paste(Error_Type, "-", Error_Description), fill = paste(Error_Type, "-", Error_Description)) +
  geom_bar(position = "dodge") +
  labs(
    title = "HL7 User Errors",
    x = "",
    y = "Number of errors",
    fill = "Error type"
  ) +
  theme_bw() + 
  geom_text(data=summary_with_desc2,aes(x=paste(Error_Type, "-", Error_Description),y=n,label=n),vjust=0.5, hjust=-0.1) +
  coord_flip() +
  scale_y_log10() +
  theme(legend.position = "none") +
  expand_limits(y=10000)

desc_plot2

# print files
write.xlsx(summary_with_desc2, paste0(DIR, "/Summary of user error descriptions.xlsx"), overwrite = TRUE)
write.xlsx(user_error_records, paste0(DIR, "/Records with user errors.xlsx"), overwrite = TRUE)
```
# Plot the errors over time
```{r}
user_error_records <- user_error_records %>% 
  mutate(
    Error_Week = paste0(month(anydate(Record_Date_Start)), "-", year(anydate(Record_Date_Start))),
    Record_Date_Start = anydate(Record_Date_Start))  

# plot
error_week_plot1 <- ggplot(user_error_records %>% filter(Error_Week!="NA-NA")) +
  aes(x = Error_Week, fill = Error_Type) +
  geom_bar(position = "dodge") +
  labs(
    title = "User Errors Over Time",
    x = "Week of record submission",
    y = "Number of errors",
    fill = "Submission_Week"
  ) +
  coord_flip() 

error_week_plot1

records <- records %>% 
  mutate(
    Error_Week = paste0(month(anydate(Record_Date_Start)), "-", year(anydate(Record_Date_Start))),
    Record_Date_Start = anydate(Record_Date_Start))  

# plot
error_week_plot2 <- ggplot(records %>% filter(Error_Week!="NA-NA")) +
  aes(x = Error_Week, fill = Error_Desc) +
  geom_bar(position = "dodge") +
  labs(
    title = "Errors Over Time",
    x = "Week of record submission",
    y = "Number of errors",
    fill = "Submission_Week"
  ) +
  coord_flip() +
  scale_y_log10() +
  expand_limits(y=10000)

error_week_plot2

```


    ----- Work with Vim Transactions -----
    
Pull the Vim records and date
```{r}
# Set up regex patterns
pat <- "\\n.*,.*-.*\\d{1,2}\\/\\d{1,2}\\/\\d{2,4}.*\\n.*\\d{1,2}\\/\\d{1,2}\\/\\d{2,4}" # pull text with the date
pat2 <- "\\d{1,2}\\/\\d{1,2}\\/\\d{2,4}" # Will match any date
pat3 <- "\\w+\\s*,\\s*\\w+" # Will match name

# pull all record dates from Vim section of PDF
vim_important <- str_locate_all(Vim, pat)

# Get a count of total records within the Vim section
f <- 0
for(i in vim_important){f <- f+(dim(i)[1])}

# Set up a table to hold the records
vim_records <- tibble(
  "DOB" = rep("NA", f), 
  "Name" = rep("NA", f),
  "Submission_Date" = rep("NA", f))

n <- 1
for(i in 1:length(Vim)[1]){
  if(i %% 10==0) {print(paste("Processing page ", i))}
  if(i %% 100==0) {cat(round((i/length(Vim))*100, digits=0), "% completed...")}  
  if(length(vim_important[[i]]==0)){} # Skip if no records on this page
  if(length(vim_important[[i]]>0)){
    for(j in 1:dim(vim_important[[i]])[1]){ # for each dob in that page
      # pull start and end index for this item
      vim_start <- vim_important[[i]][j,1]
      vim_end <- vim_important[[i]][j,2]
      # use the indices to grab the string
      V <- substr(Vim[i], vim_start, vim_end)
      # DOB will be the first date match in the string
      B <- str_extract(V, pat2)
      # Find the submission date at the end of the string
        # nest using the second pattern, pulling from the end of the string only
      D <- str_extract(substr(V, nchar(V)-11, nchar(V)), pat2)
      # Find the Name
      Nam <- str_extract(V, pat3)

      # apply the string to the records table
      vim_records$DOB[n] <- B
      vim_records$Name[n] <- Nam      
      vim_records$Submission_Date[n] <- D

      n <- n+1 # iterate up for next loop & next row in final table# pull date for this record
    }
  }
}
```


Review Vim and print to file
```{r}
# Summarize the transmissions
vim_summary <- vim_records %>% 
  mutate(
    Submission_Week = week(anydate(Submission_Date)),
    Submission_Date = anydate(Submission_Date))  %>% 
  group_by(DOB, Name, Submission_Week) %>% 
  summarise(n=n())

# plot
vim_plot <- ggplot(vim_summary) +
  aes(x = Submission_Week, fill = Submission_Week) +
  geom_bar(position = "dodge") +
  labs(
    title = "Successful Vim Transmissions",
    x = "Week of record submission",
    y = "Number of transmissions",
    fill = "Submission_Week"
  ) +
  theme_bw() + 
  theme(legend.position = "none") 

vim_plot

# write the vim transmissions to file
write.xlsx(vim_records, paste0(DIR, "/Records of successful Vim Transmissions.xlsx"), overwrite = TRUE)

```
