---
title: "Printing muliple xlsx sheets"
author: "Samantha L Bell"
date: "10/29/2020"
output: html_document
---
_____________________________________________________
This code:
1- loads the base Iris dataset from R
2- manipulates the data
3- prints multiple sheets to a single excel workbook

Load the necessary library packages
```{r packages}
library(datasets)
library(openxlsx)
library(tidyverse)
```

Set up the data & empty workbook
```{r}
# data from the datasets library
data(iris)

# an empty workbook object, to be added to and then printed to file
iris_wb <- createWorkbook() 

# Where the output will be printed 
outLocation <- "C:/Desktop"

```

Add a summary page to the workbook 
  * Remember, we are not printing the workbook yet, just adding to the object!
```{r}
addWorksheet(iris_wb, "Summary") # Add a blank worksheet and give it a name

sumIris <- summary(iris) # Save a summary table to a variable

writeData(iris_wb, "Summary", sumIris, startRow = 1, startCol = 1) # Write to the worksheet
```


Add a table to the workbook, with some text above it 
 * Adjusting the startRow or startCol allows for separate writing commands to same sheet
```{r}
addWorksheet(iris_wb, "Sepal Length")

sepalTable <- iris %>% group_by(Species) %>% summarise(Ave.Sepal.Length = mean(Sepal.Length)) # Summary of mean sepal length by species

# Write a title to the sheet first, followed by the table
writeData(iris_wb, "Sepal Length", "Table of average sepal length by species", startCol = 1, startRow = 1)
writeData(iris_wb, "Sepal Length", sepalTable, startCol = 1, startRow = 3)
```


Add another table to the workbook, but append a graph on the same sheet
  * Separate commands allow for writing data vs writing plots
```{r}
addWorksheet(iris_wb, "Petal Length")

petalTable <- iris %>% group_by(Species) %>% summarise(Ave.Petal.Length = mean(Petal.Length))

# Make a plot
petalPlot <- plot(y = as.numeric(iris$Petal.Length),               # The petal length on y axis as a continuous variable
                  x = as.factor(iris$Species),                     # The species on x axis as a factor / discrete variable
                  col = 3:6 ,                                      # 3 of the base colors available in R
                  xlab = "", ylab = "Petal Length",                # Labels for x and y axis
                  cex.axis = 0.6,                                  # Text size
                  title(main="Length of Petals", col.main="black") # Main title and color
)

writeData(iris_wb, "Petal Length", petalTable, startRow = 1, startCol = 1) # Write the table

# This command requires that the dersire plot be currently displayed in R in order to print to the sheet
insertPlot(iris_wb, "Petal Length", fileType = "png",
           width = 10, height = 6,                    # make the size of the plot
           startCol = 6)                              # Print to the right of the table by starting at column 6
```


Save the workbook to file!
```{r}
saveWorkbook(iris_wb, file = paste0(outLocation, "/", Sys.Date(), " Fun with Iris Data.xlsx"), overwrite = TRUE)

```




