---
title: "Stringr Tips"
author: "Samantha L Bell"
date: "12/3/2020"
output: html_document
---

Manipulating strings and pulling patterns of text is a frequent coding task, and can be a challenge. Among the many packages available for manipulating strings is the "stringr" package. Today's tip dives into some of the functions stringr has to offer:

1) str_locate - identifying the start and end character indices of your search text
2) word - pulling entire words based on their index or range of indices. "Words" are defined by a separator (default is a single space)
3) str_split - separating a string into a list, using a provided separator
4) str_extract - print matched text from a string

Also touched on is the use of the collapse option from paste() - allowing for quick creation of long strings with separator text. 

Ready! Set! Go! 


Load our stringr package (install if needed)
```{r loadPackages}
library(stringr)
```

_____________________________________________________________________________________________________________________
First, let's try to locate a string of text within some sentences. 

#str_locate will provide us with the start and end indices of our desired search text
  The function runs as str_locate(string, pattern), and will give the location of matched text within the string.
  To ignore the case of the text, wrap the pattern in regex() and use ignore.case = TRUE
```{r locateString}
exampleSong <- "All God's creatures got a place in the choir. Some sing low and some sing higher. Some sing out loud on a telephone wire. Some just clap their hands, or paws, or anything they've got now."

toMatch <- "some sing"

str_locate(exampleSong, toMatch) # Find the FIRST location, NOT ignoring case
str_locate(exampleSong, regex(toMatch, ignore_case = TRUE) ) # Find the FIRST location, IGNORING case
str_locate_all(exampleSong, regex(toMatch, ignore_case = TRUE) ) # Find ALL locations, IGNORING case

```
_____________________________________________________________________________________________________________________
What if we wanted to do the opposite, and capture a word or words based on their positions?

# word() will let us pull from a string based on a set position or range, with the ability to specify the separator as well
To see how this works, let's first use our example song text again.
  We can pull specific words using those positions and word(string, start, end, sep)
  When not provided, the separator is assumed to be a single space: " "
```{r word}
# Pull the first instance of "some sing"
word(exampleSong, start = 10, end = 11) # Notice that this function counts by WORD, not CHARACTER

# You can simplify the command if only providing the start position 
word(exampleSong, 16) # Pull the 16th word

# A negative number will allow you to count backwards
word(exampleSong, start = -6, end = -1) # Pull from the end of the string
```

Now that we have an understanding of how this works, we can take a look at a different application. 
  In this case, we have a string of test dates from 3 patients. 
  We want to pull certain test dates out using word()
  Since our dates are stored using .. as a separator, we must specify this when using word()
    ** Remember that "." is a regex symbol with its own meaning! To avoid this being interpreted by R, we must escape the characters:
      A double slash before a symbol will let R know to read it as a normal character. \\ is "escape"
      Therefore ".." becomes "\\.\\." when we specify the separator
      
```{r wordDates}
# Create our table with a row for each patient
testDates <- rbind( A = "03/23/2020..04/02/2020..05/14/2020..09/25/2020",
                    B = "05/04/2020..07/15/2020..08/21/2020",
                    C = "09/13/2020..09/24/2020..10/01/2020..11/25/2020" )

# Find the first test date for each patient
for(patient in 1:dim(testDates)[1]){ # for each patient row in range 1 to the number of rows
  print(word(testDates[patient], start = 1, sep = "\\.\\."))
}

### For comparison, what happens if you forget to escape the dots?
word(testDates[1], start = 1, sep = "..")

# Find the most recent test date for all patients
for(patient in 1:dim(testDates)[1]){ # for each patient row in range 1 to the number of rows
  print(word(testDates[patient], start = -1, sep = "\\.\\."))
}
```
_____________________________________________________________________________________________________________________
We can also use str_split() to separate captured ranges of dates, applying the same separator text
  - This makes them much easier to use for other purposes (graphs, calculations, etc)
  
str_split() takes in the string to be split as well as the pattern used to identify splitting locations.  

  For Patient A, we will:
  1- Find the 1st through 3rd dates
  2- Print the non-split captured text 
  3- Split the dates on the same separator used to find them, and print as a list
```{r wordDatesSplit}
# Find the 1st through 3rd tests for Patient A
cat("These are the dates not split, as a string\n")
dates <- word(testDates[1], start = 1, end = 3, sep = "\\.\\.") # assign to a variable
dates # view

# Print the split dates 
cat("\nThese are the split dates as a list")
print(str_split(dates, pattern = "\\.\\."))
```
_____________________________________________________________________________________________________________________
What if we need to capture only certain dates as a list, if we don't want a range, but need to handpick specific dates?
  Instead of using word() and str_split() together, we can specify a list of indices to the word function itself!
  
  Let's pick out test 1 and 3 for each patient.
  Using c() to provide a list of positions, we can select certain test dates and return them as a list.
  
```{r wordDatesList}
for(patient in 1:dim(testDates)[1]){
  print(word(testDates[patient], start = c(1, 3), sep = "\\.\\."))
}
```


_____________________________________________________________________________________________________________________
Finally, we will take a look at str_extract(), which prints matched text from a string. 

Let's say we have a list of hospital diagnosis codes, but we only want to pull out certain ones
The codes we are looking for start with "T" and are followed by the numbers 36-50. (ex: T36)

#We want to create a regex pattern to match against which will:
- Be one string of text (required for regex)
- Have each desired code separated by the or symbol "|", so that we pull matches to any of the possible codes in our desired list
- Not require us to type out every code by hand

```{r makePattern}
# First, we make our regex pattern text. We know we can use paste to get the list of numbers:
cat("This is just a character vector of numbers\n")
paste(36:50)

# But that is not quite what we want. 
 # Remember, to be a regex pattern, it must be one string of text. 
 # If we use the collapse option, we can get one string and can specify how the numbers are separated
    # The "|" is used as "or" in the regex, so each code must be separated by this. 
    # By adding the letter "T" to our collapse, we add the leading T to each trailing number
cat("\nThis is a collapsed string of codes with or symbols, but missing leading T\n")
paste(paste(36:50), collapse ="|T") 

# Almost there! But we are still missing the first "T"
  # This is an easy fix. We can use paste0 to quickly add the leading T without any separator
  # The paste0 code will simply surround our previous pattern: paste0("T", previousPattern)
pat <- paste0("T", paste(paste(36:50), collapse = "|T") ) 
# View our finished pattern
cat("\nThis has every code we need to match, separated by the or symbol\n")
pat
```
#Now let's find some matches!

Using the string of codes we are provided as input, the str_extract() function lets us pull out matching text by using the pattern we created in the step above
```{r matchCodes}
codes <- "T35..T42..J12..H14..M44..T55..A23..T47..A36..T43..M21..H51" # String of codes to pull matches from

str_extract(codes, pat) # Print the FIRST match
str_extract_all(codes, pat) # Print ALL matches
```


HAPPY PROGRAMMING! 


