---
title: "Calculate and Combine Vendor Data"
author: "SLB"
date: '2022-03-29'
output: html_document
---
This script takes in the raw monthly vendor data, cleans it up, makes calculations, and exports the standardized fields for each vendor. 
The processed data will be able to be added to the ongoing vendor data Excel files- used for powerbi reporting. 


##Load the libraries we need
```{r libraries}
library(tidyverse)
library(lubridate)
library(anytime)
library(openxlsx)
library(svDialogs)
library(readxl)
library(stringr)
library(openxlsx)
library(janitor)
```

##Load the raw data
# Each month will require the following:
#     COVETRUS: CR_RevByVet_Detail_NVA_[date].xls
#     VETSOURCE: [date] NVA Ext.xlsx
#     Hill's to Home: [date] NVA Recon.xlsx
#     Vetcove: [Year]NVA_[Month]_Payouts.xlsx


```{r}
# What period are we working on?
period <- dlg_input(message="Enter the date for the month period we are working with, with '01' as the day field (ex: 2022-06-01)")$res
period <- anydate(period)

# Hills
dlg_message(paste("Select the monthly Hill's to Home file for", period))
H2H_PATH <- dlg_open("N:/Category Management/Home Delivery Reporting/Raw Vendor Data/H2H Monthly")$res
H2H_SHEETS <- readxl::excel_sheets(H2H_PATH)
H2H <- readWorkbook(H2H_PATH, sheet = grep("HTH", H2H_SHEETS, ignore.case=TRUE), detectDates = TRUE)

# Covetrus
dlg_message(paste("Select the monthly Covetrus file for", period))
CVET_PATH <- dlg_open("N:/Category Management/Home Delivery Reporting/Raw Vendor Data/Covetrus Monthly")$res
CVET_SHEETS <- readxl::excel_sheets(CVET_PATH)
CVET_TIBBLES <- lapply(CVET_SHEETS, function(x) readxl::read_xls(CVET_PATH, sheet=x))
 # join all CVET sheets together
CVET <- CVET_TIBBLES[[1]] # start with the first one
colnames(CVET) <- 1:dim(CVET_TIBBLES[[1]])[2]
 for(i in 2:length(CVET_TIBBLES)){ #add others
   j <- CVET_TIBBLES[[i]]
   colnames(j) <- colnames(CVET)
   CVET <- rbind2(CVET, j)
 }

# VetSource
dlg_message(paste("Select the monthly Vetsource file for", period))
VS_PATH <- dlg_open("N:/Category Management/Home Delivery Reporting/Raw Vendor Data/VetSource Monthly")$res
VS_SHEETS <- readxl::excel_sheets(VS_PATH)
VS <- readxl::read_xlsx(VS_PATH, sheet=grep("Data", VS_SHEETS)) #Load the sheet with the Data for that month

# Vetcove
dlg_message(paste("Select the monthly Vetcove file for", period))
VC_PATH <- dlg_open("N:/Category Management/Home Delivery Reporting/Raw Vendor Data/Vetcove Monthly")$res
VC <- readxl::read_xlsx(VC_PATH)
```

# Clean up Covetrus
```{r}
# Remove header if not already the column names
headRow <- which(CVET[,1] == "Practice Name") 
colnames(CVET) <- CVET[headRow,]
CVET <- CVET[-(1:headRow),]

# Remove "total" rows
CVET <- CVET %>% mutate(totalRow = grepl("(tot)|(dvm)|(vmd)", `Practice Name`, ignore.case = TRUE))
totalRows <- CVET %>% filter(totalRow==TRUE)
CVET <- CVET %>% filter(totalRow==FALSE) %>% select(-totalRow) %>% filter(!is.na(`Cart ID`))

# How many missing hospital numbers are there?
missingHosp <- is.na(CVET$`Hospital #`)
table(missingHosp)
dlg_message(paste("You have", length(which(missingHosp==TRUE)), "Covetrus rows with missing NVA hospital number."))

# Pull in key of Covetrus site numbers to DVM names (from previous reports)
CVET_key <- readxl::read_xlsx("N:/Category Management/Home Delivery Reporting/Other Resources/Covetrus hospital number and DVM.xlsx") %>% filter(!is.na(Hospital))

# Fill in some missing site numbers
CVET <- CVET %>% mutate(rowNum = row_number())

for(i in 1:dim(CVET[missingHosp,])[1]){
  row <- CVET$rowNum[missingHosp][i]
  lookup <- CVET_key$Hospital[which(CVET_key$Veterinarian==CVET$Veterinarian[row])]
  if(!is_empty(lookup)){CVET$`Hospital #`[row] <- lookup}
}

# Check again for missing numbers
missingHosp2 <- is.na(CVET$`Hospital #`)
table(missingHosp2)
dlg_message(paste("We were able to successfully fill in", length(which(missingHosp==TRUE))-length(which(missingHosp2==TRUE)), "(", 
round((length(which(missingHosp==TRUE))-length(which(missingHosp2==TRUE)))/length(which(missingHosp==TRUE))*100, digits = 0), "% ) Covetrus rows which had a missing NVA hospital number. Be sure to keep the reference file 'Covetrus rows with missing NVA hospital number.xlsx' updated for future use."))
# Remove any "s" from hospital ids entries - these are sister sites
CVET$`Hospital #` <- gsub("\\.s", "", CVET$`Hospital #`, ignore.case=TRUE)

# Fix hospital numbers: There is no hospital 828, it should be 142
CVET$`Hospital #` <- gsub("828", 142, CVET$`Hospital #`)

# Export unique DVM and site number combos to add some to the key (if any are new)
vets <- CVET %>% group_by(`Hospital #`, `Practice Name`, Veterinarian) %>% summarise() 

missingHosp <- vets %>% filter(is.na(`Hospital #`)&!is.na(Veterinarian))
write.xlsx(missingHosp, "N:/Category Management/Home Delivery Reporting/Monthly Combined Calculations/CVET DVMs missing hospital number.xlsx")

vets <- vets %>% filter(!is.na(`Hospital #`)&!is.na(Veterinarian))
write.xlsx(vets, "N:/Category Management/Home Delivery Reporting/Monthly Combined Calculations/Use to update CVET DVM list.xlsx")

# View DVMs missing site numbers still
CVET[missingHosp2,] %>% group_by(`Hospital #`, `Practice Name`, Veterinarian) %>% summarise() 
```

## Make the new standardized fields for Covetrus
```{r}
# reformat ship date
CVET$Shipped <- openxlsx::convertToDate(CVET$Shipped)
# Make all NA numbers 0
CVET$`Product Revenue`[is.na(CVET$`Product Revenue`)] <- 0
CVET$Handling[is.na(CVET$Handling)] <- 0 
CVET$`Client Total`[is.na(CVET$`Client Total`)] <- 0
CVET$`Product Cost`[is.na(CVET$`Product Cost`)] <- 0
CVET$`Service Fee`[is.na(CVET$`Service Fee`)] <- 0
CVET$`Sales Tax`[is.na(CVET$`Sales Tax`)] <- 0
CVET$`Practice Total`[is.na(CVET$`Practice Total`)] <- 0
CVET$`Net Margin`[is.na(CVET$`Net Margin`)] <- 0

CVET_CLEAN <- CVET %>% mutate(
  `Hospital Number` = gsub("\\*failed to decode utf16\\*", "", as.character(`Hospital #`)),
  `Hospital Name` = "",
  Division = "",
  Group = "",
  `Cart Number` = as.character(`Cart ID`),
  Period = period,
  Client = `Client Name`,
  `Pet Name` = Patient,
  Description = gsub("\\*failed to decode utf16\\*", "", Description), 
  Quantity = Qty, 
  Revenue = as.numeric(`Client Total`),
  Cost = abs(as.numeric(`Product Cost`)),
  `Full Cost` =  abs(as.numeric(`Practice Total`)),
  `Net Profit` = `Net Margin`, 
  Partner = "CVET", 
  GL = as.numeric(`NVA GL#`), 
  Autoship = `Autoship?`, 
  Refill = `Refill?`, 
  Rx = Rx, 
  `Client PIMS ID` = "",
  `Patient PIMS ID` = ""
)
CVET_CLEAN <- CVET_CLEAN %>% select(`Hospital Number`, `Hospital Name`, Group, Division, `Cart Number`, Period, Client, `Pet Name`, Description, Quantity, 
                                Revenue, Cost, `Full Cost`, `Net Profit`, Partner, GL, Autoship, Refill, Rx, `Client PIMS ID`, `Patient PIMS ID`)

```
## Make the new standardized fields for VetSource
```{r}
# Remove rows without order number - these are blank or total rows
VS <- VS %>% filter(!is.na(`Order Vso ID`))

# Turn NA numbers to 0
VS$`Unit Price`[is.na(VS$`Unit Price`)] <- 0
VS$`Groups Reporting: Net Sales`[is.na(VS$`Groups Reporting: Net Sales`)] <- 0
VS$`Groups Reporting: DVM Auth/Affiliate Commission`[is.na(VS$`Groups Reporting: DVM Auth/Affiliate Commission`)] <- 0
VS$`Sales Tax`[is.na(VS$`Sales Tax`)] <- 0
VS$`Wholesale Amount`[is.na(VS$`Wholesale Amount`)] <- 0
VS$`Wholesale Tax`[is.na(VS$`Wholesale Tax`)] <- 0
VS$`Marketplace Facilitator Tax`[is.na(VS$`Marketplace Facilitator Tax`)] <- 0
VS$`Credit Card Fee`[is.na(VS$`Credit Card Fee`)] <- 0
VS$`Vetsource Fee`[is.na(VS$`Vetsource Fee`)] <- 0
VS$`Promotion Credit Amount`[is.na(VS$`Promotion Credit Amount`)] <- 0
VS$`Net Profits`[is.na(VS$`Net Profits`)] <- 0
VS$`Net Proceeds`[is.na(VS$`Net Proceeds`)] <- 0

VS_CLEAN <- VS %>% mutate(
  `Hospital Number` = as.character(`Corporate ID`),
  `Hospital Name` = "",
  Division = "",
  Group = "",
  `Cart Number` = as.character(`Order Vso ID`),
  Period = period,
  Client = paste(`Client First Name`, `Client Last Name`),
  `Pet Name` = `Patient Name`,
  Description = `Display Name`, 
  Quantity = `Line Quantity`, 
  Revenue = as.numeric(`Groups Reporting: Net Sales`)+as.numeric(`Promotion Credit Amount`),
  Cost = as.numeric(`Wholesale Amount`),
  `Full Cost` =  as.numeric(`Sales Tax`)  + abs(as.numeric(`Wholesale Amount`)) + as.numeric(`Wholesale Tax`) + as.numeric(`Marketplace Facilitator Tax`) + abs(as.numeric(`Credit Card Fee`)) + abs(as.numeric(`Vetsource Fee`)),
  `Net Profit` = `Net Profits`, 
  Partner = "VS", 
  GL = as.numeric(`GL Code Per NVA`), 
  Autoship = "Unknown", 
  Refill = "Unknown", 
  Rx = "Unknown", 
  `Client PIMS ID` = "",
  `Patient PIMS ID` = ""
)
VS_CLEAN <- VS_CLEAN %>%  select(`Hospital Number`, `Hospital Name`, Group, Division, `Cart Number`, Period, Client, `Pet Name`, Description, Quantity, 
                                Revenue, Cost, `Full Cost`, `Net Profit`, Partner, GL, Autoship, Refill, Rx, `Client PIMS ID`, `Patient PIMS ID`)
```

## Make the new standardized fields for Hills to Home
```{r}
# Remove extra rows
H2H <- H2H[,1:25]
# Remove total row by including only valid order numbers
H2H <- H2H %>% mutate(Order = PoNumber) %>% filter(grepl("HTH", `Order`))

# Turn NA numbers to 0
H2H$Order.Total[is.na(H2H$OrderTotal)] <- 0
H2H$Sales.Tax[is.na(H2H$SalesTax)] <- 0
H2H$`Shipping./.Surcharge`[is.na(H2H$`Shipp_Surcharge`)] <- 0
H2H$List.Price[is.na(H2H$ListPrice)] <- 0
H2H$Service.Fee[is.na(H2H$ServiceFee)] <- 0
H2H$Total.Discount[is.na(H2H$Total_Discount)] <- 0
H2H$Clinic.Funded.Discount[is.na(H2H$Promo_Discount)] <- 0
H2H$Clinic.Sales.Tax[is.na(H2H$Clinic_Sales_Tax)] <- 0
H2H$Profit[is.na(H2H$VProfit)] <- 0
H2H$Stripe.Fee[is.na(H2H$StripeFee)] <- 0
H2H$Net[is.na(H2H$VNet)] <- 0
H2H$Price[is.na(H2H$PPrice)] <- 0

# We need to make sure that all refund rows are negated in the proper columns, if not, we will fix them. 
	#Refund rows should have a negative value for: VNet, VProfit, Total_Discount, OrderTotal
	#Refund rows should have a positive value for: SalesTax, ListPrice, ServiceFee, StripeFee
H2H_refunds <- H2H %>% filter(Refund == "Refund")
if(sum(H2H_refunds$VNet)>0){
  dlg_message("The Hills to Home data does not have refund rows as negative rows - this will be fixed and replaced.")
  H2H_refunds <- H2H_refunds %>% mutate(
    VNet = -abs(VNet), 
    VProfit = -abs(VProfit), 
    Total_Discount = -abs(Total_Discount), 
    OrderTotal = -abs(OrderTotal), 
    SalesTax = abs(SalesTax), 
    ListPrice = abs(ListPrice), 
    ServiceFee = abs(ServiceFee), 
    StripeFee = abs(StripeFee)
  )
  H2H_sales <- anti_join(H2H, H2H_refunds)
  H2H <- rbind(H2H_sales, H2H_refunds)
  if(sum(H2H_refunds$VNet)<0){
  dlg_message("The Hills to Home data now has refund rows as negative rows. Replacement rows have been saved.")
  }
}else(dlg_message("The Hills to Home data already has refund rows as negative rows - no action will be taken."))



H2H_CLEAN <- H2H %>% mutate(
  `Hospital Number` = as.character(hosp.number), 
  `Hospital Name` = "", 
  Group = "",
  Division = "",
  `Cart Number` = as.character(Order), 
  Period = period,
  Client = PetParent,
  `Pet Name` = "",
  Description = "", 
  Quantity = 1,
  Revenue = as.numeric(OrderTotal)+as.numeric(SalesTax)+ abs(as.numeric(Promo_Discount))+as.numeric(Clinic_Sales_Tax), #UPDATED 5/2023 because Promo Discount field is negative and already subtracted from OrderTotal
  Cost = abs(as.numeric(ListPrice)+ as.numeric(Total_Discount)), # list price is neg, discount is pos, then abs reverses the overall sign at end
  `Full Cost` =  abs(as.numeric(ListPrice)+ as.numeric(Total_Discount)) + abs(as.numeric(SalesTax)) + abs(as.numeric(ServiceFee)) + abs(as.numeric(StripeFee)),
  `Net Profit` = VNet, 
  Partner = "H2H",
  GL = 14, 
  Autoship = "Unknown", 
  Refill = "Unknown", 
  Rx = "Unknown", 
  `Client PIMS ID` = "",
  `Patient PIMS ID` = ""
)
H2H_CLEAN <- H2H_CLEAN %>%  select(`Hospital Number`, `Hospital Name`, Group, Division, `Cart Number`, Period, Client, `Pet Name`, Description, Quantity, 
                                Revenue, Cost, `Full Cost`, `Net Profit`, Partner, GL, Autoship, Refill, Rx, `Client PIMS ID`, `Patient PIMS ID`)
```

## Clean up and make the new standardized fields for VetCove
```{r}
# Remove header
headRow <- which(VC[,2] == "Stripe Payout ID")
colnames(VC) <- VC[headRow,]
VC <- VC[-(1:headRow),]

# Pull in NVA ids
VC_dir <- readxl::read_excel("N:/Category Management/Home Delivery Reporting/FILES FOR WORKBOOKS/Vetcove Facility Directory.xlsx") %>% mutate(`Vetcove ID` = as.character(`Vetcove ID`))
VC <- VC %>% left_join(., VC_dir, by = c("Clinic ID" = "Vetcove ID"))

# Pull in NVA GL codes and map the items
VC_GL <- readxl::read_excel("N:/Category Management/Home Delivery Reporting/FILES FOR WORKBOOKS/Vetcove GL Mapping.xlsx")

# Look for a match in the main and sub categories both - then pull NVA GL code from key
VC$NVA_GL <- ""
for(i in 1:dim(VC)[1]){
 VC$NVA_GL[i] <- VC_GL$`NVA GL Code`[which(VC_GL$`Vetcove Main Category` %in% str_split(VC$Categories[i], ", ", simplify = TRUE)[1:3] & 
                   VC_GL$`Vetcove Sub Category` %in% str_split(VC$Categories[i], ", ", simplify = TRUE)[1:3])][1]
}

# Make NA nunbers into 0
VC$`Retail Sale Price`[is.na(VC$`Retail Sale Price`)] <- 0
VC$`Urgency Fee`[is.na(VC$`Urgency Fee`)] <- 0
VC$`Handling Fee`[is.na(VC$`Handling Fee`)] <- 0
VC$`Shipping Charged`[is.na(VC$`Shipping Charged`)] <- 0
VC$`Refunded Amount`[is.na(VC$`Sponsored Refund Amount`)] <- 0
VC$`Product Cost`[is.na(VC$`Product Cost`)] <- 0
VC$`Vendor Fees`[is.na(VC$`Vendor Fees`)] <- 0
VC$`Clinic Discounts`[is.na(VC$`Clinic Discounts`)] <- 0
VC$`Processing Fees`[is.na(VC$`Processing Fees`)] <- 0
VC$`Net Revenue`[is.na(VC$`Net Revenue`)] <- 0


# Check for missing information
a <- VC %>% filter(is.na(`NVA ID`)) %>% group_by(`Clinic Name`) %>% summarise()
b <- VC %>% filter(is.na(`NVA_GL`)) %>% group_by(Categories) %>% summarise()

if(dim(a)[1]>0){stop("You have missing Vetcove hospital IDs! Please check table a and update your Vetcove Facility Directory accordingly. Rerun the Vetcove code block once updates to file is complete.")}
if(dim(b)[1]>0){stop("You have missing Vetcove GL codes! Please check table b and update your Vetcove GL Mapping file accordingly. Rerun the Vetcove code block once updates to file is complete.")}

# Note - Rebate discounts are not in these calculations because they are already represented in what the retail sale price for the client was)
# Clean up other fields
VC_CLEAN <- VC %>% mutate(
  `Hospital Number` =  `NVA ID`, 
  `Hospital Name` = `Clinic Name`, 
  Group = "",
  Division = "",
  `Cart Number` =  `Order No.`, 
  Period = period,
  Client =  `Client Name`,
  `Pet Name` = `Patient Name`,
  Description = `Item Name`, 
  Quantity =  Quantity,
  Revenue =  as.numeric(`Retail Sale Price`) + abs(as.numeric(`Urgency Fee`)) + abs(as.numeric(`Handling Fee`)) + abs(as.numeric(`Shipping Charged`))
	- abs(as.numeric(`Sponsored Refund Amount`)),
  Cost =  abs(as.numeric(`Product Cost`)),
  `Full Cost` =  abs(as.numeric(`Product Cost`) + as.numeric(`Vendor Fees`) + as.numeric(`Clinic Discounts`)+ as.numeric(`Processing Fees`)),
  `Net Profit` = `Net Revenue`,
  Partner = "Vetcove",
  GL = NVA_GL, 
  Autoship = Autoship, 
  Refill = "Unknown", 
  Rx = "Unknown", 
  `Client PIMS ID` = `Client PIM ID`,
  `Patient PIMS ID` = `Patient PIM ID`
)
VC_CLEAN <- VC_CLEAN %>%  select(`Hospital Number`, `Hospital Name`, Group, Division, `Cart Number`, Period, Client, `Pet Name`, Description, Quantity, 
                                Revenue, Cost, `Full Cost`, `Net Profit`, Partner, GL, Autoship, Refill, Rx, `Client PIMS ID`, `Patient PIMS ID`)


```

## Combine all calculated fields together
```{r}
CVET_CLEAN <- CVET_CLEAN %>% mutate_all(as.character)
VS_CLEAN <- VS_CLEAN %>% mutate_all(as.character)
H2H_CLEAN <- H2H_CLEAN %>% mutate_all(as.character)
VC_CLEAN <- VC_CLEAN %>% mutate_all(as.character)

# Warning if a table is empty
if(dim(CVET_CLEAN)[1]==0 | dim(VS_CLEAN)[1]==0 | dim(VC_CLEAN)[1]==0 | dim(H2H_CLEAN)[1]==0){stop("One of your vendor tables is empty, please repair and rerun before combining the tables!")}
ALL <- full_join(CVET_CLEAN, VS_CLEAN) %>% full_join(H2H_CLEAN) %>% full_join(VC_CLEAN) 
ALL$GL[is.na(ALL$GL)] <- 0
ALL$`Hospital Number`[is.na(ALL$`Hospital Number`)] <- 0
```

## Print to file to allow for QC and manual addition to ongoing file(s)
```{r}
# Make visits summary
t <- ALL %>% group_by(`Hospital Number`, Partner) %>% summarise(Revenue = sum(as.numeric(Revenue), na.rm=TRUE), Visits = length(unique(`Cart Number`)))
t <- adorn_totals(t)
colnames(t) <- c("Hospcode", "Vendor","OrderAmt", "OrderQty")

# Print files
write.xlsx(ALL, paste0("N:/Category Management/Home Delivery Reporting/Monthly Combined Calculations/", period, " Monthly Vendor Calculated Fields.xlsx"))

write.xlsx(t, paste0("N:/Category Management/Home Delivery Reporting/Monthly Combined Calculations/", period, " HD Visits.xlsx"))

dlg_message("The the visits report is ready! Please email to Gary Glover.", type="ok")

# Print some information
t
```

